---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
rawData <- read.csv("activity.csv")
data <- rawData[complete.cases(rawData),]
```

## What is mean total number of steps taken per day?

```{r}
hist(sapply(levels(data$date), function(l) {sum(subset(data, date == l)[,1])} ), main = "", xlab = "Steps")
mean(sapply(levels(data$date), function(l) {sum(subset(data, date == l)[,1])} ))
median(sapply(levels(data$date), function(l) {sum(subset(data, date == l)[,1])} ))
```

## What is the average daily activity pattern?

```{r}
sumPerInterval <- sapply(unique(data$interval), function(i) {sum(subset(data, interval == i)[,1])})
plot(unique(data$interval), sumPerInterval, type = "l")

which(sumPerInterval == max(sumPerInterval))
```

## Imputing missing values

```{r}
sum(!complete.cases(rawData))
```

Describe and show with code a strategy for imputing missing data:
Using mean for the day to fill in missing values as suggested in assigment.

```{r}
missingData <- rawData[!complete.cases(rawData),]
x <- sapply(levels(missingData$date), function(l) {mean(subset(data, date == l)[,1])} )
x[is.nan(x)] <- 0
library(sqldf)
for(i in seq(1, length(x))) {
  name <- names(x[i]);
  step <- x[i];
  rawData <- fn$sqldf(c("update rawData set steps = $step where date like '$name' and steps is null", "select * from main.rawData"));
}
hist(sapply(levels(rawData$date), function(l) {sum(subset(rawData, date == l)[,1])} ), main = "", xlab = "Steps")
mean(sapply(levels(rawData$date), function(l) {sum(subset(rawData, date == l)[,1])} ))
median(sapply(levels(rawData$date), function(l) {sum(subset(rawData, date == l)[,1])} ))

```
What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
sumPerInterval <- sapply(unique(rawData$interval), function(i) {sum(subset(rawData, interval == i)[,1])})
plot(unique(rawData$interval), sumPerInterval, type = "l")

which(sumPerInterval == max(sumPerInterval))

sum(data$steps)
sum(rawData$steps)

```
As you can notice above the method for filling NA values did not impact the data cause whole days were filled with NAs and the sum of it was 0.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
rawData[,2] <- as.Date(rawData[,2], "%Y-%m-%d")
rawData$weekday <- weekdays(rawData[,2])
weekdays <- unique(weekdays(rawData[,2]))
rawDataInWeekDay <- subset(rawData, weekday == weekdays[1] | weekday == weekdays[2] | weekday == weekdays[3] | weekday == weekdays[4] | weekday == weekdays[5])
rawDataInWeekend <- subset(rawData, weekday == weekdays[6] | weekday == weekdays[6])

par(mfrow = c (2,1), oma=c(0,1,0,0))
par(mar=c(0,3,4,4))
plot(rawDataInWeekend$interval, rawDataInWeekend$steps, type = "l", ylab = "", xlab = "", col = "blue", xaxt = "n", yaxt = "n")
axis(2, at = NULL, labels = FALSE); axis(3, at = NULL, labels = FALSE); axis(4);
par(mar=c(4,3,0,4))
plot(rawDataInWeekend$interval, rawDataInWeekend$steps, type = "l", ylab = "", xlab = "Interval", col = "blue")
axis(1); axis(4, at = NULL, labels = FALSE);
mtext("Number of steps", 2, outer = TRUE)

```